\documentclass[11pt]{letter}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\signature{Th\'eo Hermann\\Senseable City Lab\\Massachusetts Institute of Technology\\77 Massachusetts Avenue, Cambridge, MA, USA\\thermann@mit.edu}

\address{Th\'eo Hermann\\Senseable City Lab\\Massachusetts Institute of Technology\\77 Massachusetts Avenue\\Cambridge, MA, USA}

\date{\today}

\begin{document}

\begin{letter}{The Editors\\Nature Cities}

\opening{Dear Editors,}

We are pleased to submit our manuscript entitled ``Mapping Leaf Area Index (LAI) in Urban Forest'' for consideration as an Article in \textit{Nature Cities}.

As 2024 marked the warmest year on record and urban populations continue to grow, cities face an urgent need for evidence-based strategies to combat the Urban Heat Island effect, which elevates local temperatures 2--5\textdegree C above surrounding areas and exacerbates heat-related mortality. Urban trees are among the most effective countermeasures, yet quantifying their cooling potential at city scale remains a critical barrier: traditional arborist-led inventories are labor-intensive, infrequently updated, and systematically exclude private-property trees---often the majority of the urban forest. We believe this work directly addresses \textit{Nature Cities}' mission to advance interdisciplinary urban science that informs policy and planning.

Our manuscript presents an automated framework for mapping Leaf Area Index (LAI)---the key metric governing canopy shading capacity---at individual-tree resolution across entire cities. The methodology integrates airborne LiDAR point clouds, high-resolution aerial imagery (RGB and near-infrared), and machine learning (XGBoost) to produce city-wide LAI maps from standard municipal remote sensing data. Key findings include:

\begin{itemize}
    \item A predictive accuracy of R$^2$ = 0.784, representing a 39\% improvement over traditional allometric baselines, with errors comparable to manual hemispherical photography measurements.
    \item Detection of three times more trees than the existing municipal database in Amsterdam, demonstrating the framework's ability to capture the full urban forest including private-property trees.
    \item A scalable, cost-effective pipeline adaptable to variable LiDAR point densities (5--20 pts/m$^2$), enabling deployment in cities with diverse data availability.
\end{itemize}

By providing quantitative shade assessments that extend beyond public land, this framework equips municipalities with actionable intelligence for strategic tree planting in heat-vulnerable neighborhoods---bridging urban forestry, climate adaptation, and environmental justice. The hybrid approach is designed for operational scalability: processing Amsterdam's study area required only standard computing hardware, and the methodology transfers directly to any city with aerial survey programs.

This manuscript has not been published, is not under consideration, and is not in press at any other journal. There have been no prior discussions with a \textit{Nature Cities} editor regarding this work. All co-authors---Michiel van Selm, Titus Venverloo, Fabio Duarte, and Carlo Ratti---have reviewed and approved the manuscript for submission.

We believe this work will be of broad interest to \textit{Nature Cities}' readership spanning urban ecology, climate science, remote sensing, and municipal planning. We look forward to your consideration.

\closing{Sincerely,}

\end{letter}
\end{document}
