
\backmatter

% Use \bmhead* (unnumbered, no auto-bookmark) with explicit \pdfbookmark[1]
% to avoid hyperref bookmark-level-jump warnings (\bmhead = subparagraph level 5;
% previous heading is section level 1, causing level-difference warnings).
\phantomsection
\pdfbookmark[1]{Acknowledgements}{acknowledgements}%
\bmhead*{Acknowledgements}

Not applicable.

\phantomsection
\pdfbookmark[1]{Declarations}{declarations}%
\bmhead*{Declarations}

\noindent\textbf{Funding}
Not applicable.

\noindent\textbf{Competing Interests}
The authors declare no competing interests.

\noindent\textbf{Data Availability}
The Amsterdam urban tree LAI dataset is publicly available on Hugging Face at \url{https://huggingface.co/datasets/Theozinh0/LAI_UrbanTrees_Amsterdam}. Municipal tree inventory data were obtained from the City of Amsterdam open data portal.

\noindent\textbf{Code Availability}
All code and trained models are publicly available at \url{https://github.com/thrmnn/LAI}.

\noindent\textbf{Author Contributions}
T.H.\ conceptualized the study, developed the methodology and software, performed the formal analysis and investigation, curated the data, wrote the original draft, and created all visualizations. M.v.S.\ contributed to the investigation and provided writing -- review \& editing. T.V.\ supervised the research locally and provided resources, facilitating data access at the Amsterdam Intelligence Lab. F.D.\ contributed to conceptualization and supervision, and provided writing -- review \& editing. C.R.\ provided supervision, resources, funding acquisition, and project administration. All authors reviewed and approved the final manuscript.
